Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
SCP changes- results

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each flux with sufficient data given in the Supplement (Figs. S1-S15).", out.width='100%', fig.pos='H'}

@@

**check this with final data (based on Table 1):**
There were sufficient data to assess mature forest biome differences for **9** stock variables, and significant differences among biomes were detected for **6** variables (**$B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$, $DW_{tot}$**; Table 1).
C stocks had less consistent patterns across biomes (Figs. 7, S16-S26).
In **four of the six** cases (**$B_{tot}$, $B_{ag-wood}$,  $B_{root-coarse}$, $DW_{tot}$**), temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest, with tropical and temperate broadleaf forests  in between--most commonly being significantly different from temperate conifer but not boreal forests.
For $B_{ag}$, which had by far the highest sample size, tropical forests exceeded temperate conifer forests (but not significantly). 
For $B_{foliage}$,temperate broadleaf forests were lowest (again, not significantly).
The high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each stock with sufficient data given in the Supplement (Figs. S16-S26).", out.width='100%', fig.pos='H'}

@@

Average C cycles for forests <100 years old are presented in Figures 8-11.  
Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).

*ForC* contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**).
Of these, *9* increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.

@@

@@

NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)

>>>>

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction. The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each flux with sufficient data are given in the Supplement (Figs. S1-S15).", out.width='100%', fig.pos='H'}

@@

**check this with final data (based on Table 1):**
There were sufficient data to assess mature forest biome differences for **9** stock variables, and significant differences among biomes were detected for **6** variables (**$B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$, $DW_{tot}$**; Table 1).
C stocks had less consistent patterns across biomes (Figs. 7, S16-S26).
In **four of the six** cases (**$B_{tot}$, $B_{ag-wood}$,  $B_{root-coarse}$, $DW_{tot}$**), temperate conifer forests had significantly higher stocks than the other biomes, and boreal forests in the lowest, with tropical and temperate broadleaf forests in between.
For $B_{ag}$, which had by far the highest sample size, tropical forests exceeded temperate conifer forests (but not significantly). 
For $B_{foliage}$,temperate broadleaf forests were lowest (again, not significantly).
The high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.

```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). In each panel, the left scatterplot  shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction. The boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each stock with sufficient data are given in the Supplement (Figs. S16-S26).", out.width='100%', fig.pos='H'}

@@

Average C cycles for forests <100 years old are presented in Figures 8-11.  
Both C stocks and fluxes commonly increased signficantly with stand age (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).

*ForC* contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**).
Of these, *9* increased significantly with age: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.

@@

(discuss NEP well- intuition for some would be that NEP is higher in tropics. Note biome differences for young mut not mature forests, link to latitudinal differences in C sequestration potential (Anderson 2006, @cook-patton_mapping_2020)

@@

(discuss NEP well, including)
NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)
